---
title: "Reproducible Research: Peer Assessment 1"
author: "Alex Wu"
date: "August 6, 2015"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

- set working directory
- load "activity.csv" using read.csv
- view activity object

```{r, echo=TRUE}
setwd("~/Google Drive/code/r/reproducible")
activity <- read.csv("activity.csv", header = TRUE, sep = ",", quote = "\"")
activity_no_na <- na.omit(activity)
# View(activity)
# View(activity_no_na)
```


## What is mean total number of steps taken per day?

- Make a histogram of the total number of steps taken each day
    - calculate total number of steps per day
    - plot histogram

```{r, echo=TRUE}

steps_per_day_total <- aggregate(steps~date, FUN=sum, data=activity_no_na)
steps_per_day_total

hist(steps_per_day_total$steps, xlab="Steps per Day", main="Historgram of Steps per Day")


```


- Calculate and report the **mean** and **median** total number of steps taken per day



The mean total number of steps taken per day is:
```{r, echo=TRUE}
steps_per_day_total_mean <- mean(steps_per_day_total$steps)
steps_per_day_total_mean
```


The median total number of steps taken per day is:
```{r, echo=TRUE}
steps_per_day_total_median <- median(steps_per_day_total$steps)
steps_per_day_total_median
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
library(ggplot2) 
g <- ggplot(activity_no_na, aes(interval, date, color=mean(steps)))
g+geom_line()+labs(title="Average number of steps for a given interval across all days", x="Interval", y="Date")

```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
